\documentclass[10pt,a5paper]{memoir}

\usepackage[margin=0.75in]{geometry}
\input{header.tex}
\usepackage[urw-garamond]{mathdesign}

\begin{document}
\title{Algebraic Topology}
\author{Taught by Haynes Miller;\\
Notes by Sanath Devalapurkar;\\
Edited by Sanath Devalapurkar, John Ni, and others %add other folks who've
%contributed here
}
%\date{Fall 2016 -- Spring 2017}

\frontmatter

\maketitle
\input{preface.tex}
\newpage
\tableofcontents
\newpage

\mainmatter

%The current division of labor is: John'll "work on the first part (like, most
%of 905)", and skd will start off by editing/rewriting the whole of 906.

\part{18.905: an introduction to algebraic topology}\label{905}

\chapter{Homology and CW-complexes}
\input{905/lec-1-intro.tex}
\input{905/lec-2-simplices.tex}
\input{905/lec-3-categories.tex}\label{lec:3-categories}
\input{905/lec-4-more-on-categories.tex}
\input{905/lec-5-homotopy.tex}
\input{905/lec-6-homotopy-invariance.tex}\label{lec:6-homotopy-invariance-of-homology}
\input{905/lec-7-eilenberg-zilber.tex}
\input{905/lec-8-relative-homology.tex}
\input{905/lec-9-lexseq.tex}
\input{905/lec-10-eilenberg-steenrod.tex}
\input{905/lec-11-locality-principle.tex}
\input{905/lec-12-mayer-vietoris.tex}
\input{905/lec-13-locality-end-ish.tex}
\input{905/lec-14-ending-locality-and-cw-complexes.tex}
\input{905/lec-15-cw-complexes-cellular-homology.tex}
\input{905/lec-16-homology-cw-complexes.tex}
\input{905/lec-17-RPn.tex}
\input{905/lec-18-euler-char.tex}
\input{905/lec-19-eulers-thm.tex}
\input{905/lec-20-tensor-products.tex}
\input{905/lec-21-tensor-and-tor.tex}
\input{905/lec-22-more-on-tor.tex}
\input{905/lec-23-direct-limits.tex}
\input{905/lec-24-uct.tex}
\input{905/lec-25-kunneth-eilenberg-zilber.tex}

\chapter{Cohomology and duality}
\input{905/lec-26-cohomology.tex}
\input{905/lec-27-ext-cup-product.tex}
\input{905/lec-28-uct-products-in-cohomology.tex}
\input{905/lec-29-cup-products-contd.tex}
\input{905/lec-30-surfaces-bilinear-forms.tex}
\input{905/lec-31-plethora-of-products.tex}
\input{905/lec-32-cap-product-cech-cohomology.tex}
\input{905/lec-33-fully-relative-cap-product.tex}
\input{905/lec-34-cech-cohomology.tex}
\input{905/lec-35-topological-manifolds.tex}
\input{905/lec-36-fundamental-classes.tex}
\input{905/lec-37-covering-spaces-poincare.tex}
\input{905/lec-38-the-end-applications.tex}

\part{18.906 -- homotopy theory}\label{906}

\section*{Introduction}
Here is an overview of this part of the book.
\begin{enumerate}
    \item \textbf{General homotopy theory.} This includes category theory;
	because it started as a part of algebraic topology, we'll speak freely
	about it here.  We'll also cover the general theory of homotopy groups,
	long exact sequences, and obstruction theory.
    \item \textbf{Bundles.} One of the major themes of this part of the book is
	the use of bundles to understand spaces.  This will include the theory
	of classifying spaces; later, we will touch upon connections with
	cohomology.
    \item \textbf{Spectral sequences.} It is impossible to describe everything
	about spectral sequences in the duration of a single course, so we will
	focus on a special (and important) example: the Serre spectral
	sequence.  As a consequence, we will derive some homotopy-theoretic
	applications.  For instance, we will relate homotopy and homology (via
	the Hurewicz theorem, Whitehead's theorem, and ``local'' versions like
	Serre's mod C theory).
    \item \textbf{Characteristic classes.} This relates the geometric theory of
	bundles to algebraic constructions like cohomology described earlier in
	the book.  We will discuss many examples of characteristic classes,
	including the Thom, Euler, Chern, and Stiefel-Whitney classes.  This
	will allow us to apply a lot of the theory we built up to geometry.
    %\item Time permitting, there's a beautiful story that comes out (on
	%cobordism, etc).
\end{enumerate}

\chapter{Homotopy groups}
\todo[inline]{Insert an outline of the content of each lecture.}
\input{906/lec-39-limits-colimits-adjunctions.tex}
\input{906/lec-40-compactly-generated.tex}
\input{906/lec-41-basepoints.tex}
\input{906/lec-42-fiber-bundles.tex}
\input{906/lec-43-fibrations-cofibrations.tex}
\input{906/lec-44-homotopy-fibers.tex}
\input{906/lec-45-barratt-puppe.tex}
\input{906/lec-46-relative-homotopy.tex}
\input{906/lec-47-pi_1-action-hurewicz.tex}
\input{906/lec-48-examples-cw-complexes.tex}
\input{906/lec-49-relative-hurewicz.tex}
\input{906/lec-50-cellular-approximation.tex}
\input{906/lec-51-and-all-the-rest.tex}

\chapter{Vector bundles}
\input{906/lec-52-vector-bundles.tex}
\input{906/lec-53-principal-bundles.tex}
\input{906/lec-54-I-invariance.tex}
\input{906/lec-55-grassmann-model-classifying-spaces.tex}
\input{906/lec-56-ssets.tex}
\input{906/lec-57-recall.tex}
\input{906/lec-58-bundles-classifying-spaces.tex}
\input{906/lec-59-classifying-spaces.tex}

\chapter{Spectral sequences}
\input{906/lec-60-why-sseqs.tex}
\input{906/lec-61-serre-sseq.tex}
\input{906/lec-62-exact-couples.tex}
\input{906/lec-63-sseq-applications.tex}
\input{906/lec-64-edge-homomorphisms.tex}
\input{906/lec-65-serre-classes.tex}
\input{906/lec-66-mod-C-hurewicz.tex}
\input{906/lec-67-dress-sseq.tex}
\input{906/lec-68-leray-hirsch.tex}
\input{906/lec-69-integration.tex}

\chapter{Characteristic classes}
\input{906/lec-70-grothendieck-chern-classes.tex}
\input{906/lec-71-homology-of-BUn.tex}
\input{906/lec-72-chern-roots.tex}
\input{906/lec-73-steifel-whitney-classes.tex}
\input{906/lec-74-oriented-bundles.tex}

\backmatter

\nocite{*}
\bibliographystyle{alpha}
\bibliography{main}

\end{document}
