\cleardoublepage
\section*{Preface}
These are notes from the 18.905-906 sequence, taught by Haynes Miller in the
academic year 2016-2017. The goal was to tell the story of classical algebraic
topology, starting from the basics of homology and cohomology, all the way up
to the Serre spectral sequence and characteristic classes. I tried to live\TeX
the entire course. As such, there are multiple typos and mathematical errors,
which I am terribly ashamed of.

Haynes, a few other students from the class, and I are working to clean (and,
in extreme cases, rewrite) the exposition in some chapters. Hopefully this
project will be completed by the first month of 2018, in time for the second
run of 18.906.

I'd appreciate any input on where these notes can be fixed. The github
repository for the project is at
\begin{center}
    \url{https://github.com/sanathdevalapurkar/algtop-notes}.
\end{center}
Any comments can be sent to \verb|sanathd@mit.edu|.

--- Sanath

\newpage
%Here are hyperlinks to the notes from each of the semesters.
%    \begin{enumerate}
%        \item The notes for 18.905 start at section \ref{905}.
%        \item The notes for 18.906 start at section \ref{906}.
%    \end{enumerate}

%\section*{Stuff to fix}
%\begin{itemize}
%    \item The original version of the notes used
%	\verb|\cc| to denote both a script C and the complex numbers.
%	Now, \verb|\cc| denotes a script C and \verb|\cC| denotes the complex
%	numbers.
%	Everytime you find this, edit it!
%    \item Be consistent about your use of $\Hom_\cc(C,D)$ and $\cc(C,D)$ for hom-sets.
%\end{itemize}

\section*{Things to fix}
\begin{itemize}
    \item As of \today, all of Part II has been edited, with the exception of
	\S \ref{basepoints}, \S \ref{section-serre-classes}, \S
	\ref{mod-c-hurewicz}, \S \ref{dress-sseq}, \S \ref{leray-hirsch}, and
	\S \ref{gysin-sequence}.
    \item Part I remains to be edited.
    \item The original version of the notes used \verb|\cc| to denote both a
	script C and the complex numbers.  Now, \verb|\cc| denotes a script C
	and \verb|\cC| denotes the complex numbers. This is a problem that
	needs to be fixed everywhere.
\end{itemize}
