\section{Conclusions from obstruction theory}
The main result of obstruction theory, as discussed in the previous section, is
the following.
\begin{theorem}[Obstruction theory]
    Let $(X,A)$ be a relative CW-complex, and $Y$ a simple space. The map
    $[X,Y]\to [A,Y]$ is:
    \begin{enumerate}
	\item is onto if $H^n(X,A;\pi_{n-1}(Y)) = 0$ for all $n\geq 2$.
	\item is one-to-one if $H^n(X,A;\pi_n(Y)) = 0$ for all $n\geq 1$.
    \end{enumerate}
\end{theorem}
\begin{remark}
    The first statement implies the second. Indeed, suppose we have two maps
    $g_0,g_1:X\to Y$ and a homotopy $h:g_0|_{A}\simeq g_0|_{A}$. Assume the
    first statement. Consider the relative CW-complex $(X\times I,A\times I\cup
    X\times\partial I)$. Because $(X,A)$ is a relative CW-complex, the map
    $A\hookrightarrow X$ is a cofibration; this implies that the map $A\times
    I\cup X\times\partial I\to X\times I$ is also a cofibration.
    $$H^n(X\times I,A\times I\cup X\times\partial I;\pi)\simeq
    \widetilde{H}^n(X\times I/(A\times I\cup X\times\partial I);\pi) =
    H^n(\Sigma X/A;\pi)\simeq \widetilde{H}^{n-1}(X/A;\pi).$$
    %More precisely, if I have $X_{n-1}\to Y$, we get $\theta\in
    %Z^n(X,A;\pi_{n-1}(Y))$, constructed by looking at the attaching map
    %$f_\alpha$ of some $\alpha\in\Sigma_n$, to define $\theta(g)$ via
    %$\theta(g)(\alpha) = [g\circ f_\alpha]$. This captures the obstruction to
    %extending $g$ over $\alpha$. We found that $d\theta = 0$.
\end{remark}
We proved the following statement in the previous section.
\begin{prop}
    Suppose $g:X_{n-1}\to Y$ is a map from the $(n-1)$-skeleton of $X$ to $Y$.
    Then $g|_{X_{n-2}}$ extends to $X_n\to Y$ iff $[\theta(g)] = 0$ in
    $H^n(X,A;\pi_{n-1}(Y))$.
\end{prop}
An immediate consequence is the following.
\begin{theorem}[CW-approximation]
    Any space admits a weak equivalence from a CW-complex.
\end{theorem}
This tells us that studying CW-complexes is not very restrictive, if we work up
to weak equivalence.

It is easy to see that if $W$ is a CW-complex and $f:X\to Y$ is a weak
equivalence, then $[W,X]\xrightarrow{\simeq}[W,Y]$. We can now finally conclude
the result of Theorem \ref{weakhtpyequiv}:
\begin{corollary}
    Let $X$ and $Y$ be CW-complexes. Then a weak equivalence $f:X\to Y$ is a
    homotopy equivalence.
\end{corollary}
\subsection{Postnikov and Whitehead towers}
Let $X$ be path connected. There is a space $X_{\leq n}$, and a map $X\to
X_{\leq n}$ such that $\pi_i(X_{\geq n}) = 0$ for $i>n$, and
$\pi_i(X)\xrightarrow{\simeq}\pi_i(X_{\leq n})$ for $i\leq n$. This pair
$(X,X_{\leq n})$ is essentially unique up to homotopy; the space $X_{\leq n}$
is called the \emph{$n$th Postnikov section} of $X$. Since Postnikov sections
have ``simpler'' homotopy groups, we can try to understand $X$ by studying each
of its Postnikov sections individually, and then gluing all the data together.

Suppose $A$ is some abelian group. We saw, in the first part\todo{provide a
link} that there is a space $M(A,n)$ with homology given by:
\begin{equation*}
    \widetilde{H}_i(M(A,n)) = \begin{cases}
	A & i = n\\
	0 & i\neq n.
    \end{cases}
\end{equation*}
This space was constructed from a free resolution $0\to F_1\to F_0\to A\to 0$
of $A$. We can construct a map $\bigvee S^n\to \bigvee S^n$ which realizes the
first two maps; coning this off gets $M(A,n)$. By Hurewicz, we have:
\begin{equation*}
    \pi_i(M(A,n)) = \begin{cases}
	0 & i<n\\
	A & i = n\\
	?? & i>n
    \end{cases}
\end{equation*}
It follows that, when we look at the $n$th Postnikov section of $M(A,n)$, we
have:
\begin{equation*}
    \pi_i(M(A,n)_{\leq n}) = \begin{cases}
	A & i = n\\
	0 & i\neq n.
    \end{cases}
\end{equation*}
In some sense, therefore, this Postnikov section is a ``designer homotopy
type''. It deserves a special name: $M(A,n)_{\leq n}$ is called an
\emph{Eilenberg-MacLane space}, and is denoted $K(A,n)$. By the fiber sequence
$\Omega X\to PX\to X$ with $PX\simeq \ast$, we find that $\Omega K(\pi,n)\simeq
K(\pi,n-1)$. Eilenberg-MacLane spaces are unique up to homotopy.

Note that $n=1$, $A$ does not have to be abelian, but you can still construct
$K(A,1)$. This is called the \emph{classifying space} of $G$; such spaces will
be discussed in more detail in the next chapter. Examples are in abundance: if
$\Sigma$ is a closed surface that is not $S^2$ or $\RR^2$, then $\Sigma \simeq
K(\pi_1(\Sigma),1)$. Perhaps simpler is the identification $S^1\simeq K(\Z,1)$.

\begin{example}
    We can identify $K(\Z,2)$ as $\CP^\infty$. To see this, observe that we
    have a fiber sequence $S^1\to S^{2n+1}\to \CP^n$. The long exact sequence
    in homotopy tells us that the homotopy groups of $\CP^n$ are the same as
    the homotopy groups of $S^1$, until $\pi_\ast S^{2n+1}$ starts to
    interfere. As $n$ grows, we obtain a fibration $S^1\to S^\infty\to
    \CP^\infty$. Since $S^\infty$ is weakly contractible (it has no nonzero
    homotopy groups), we get the desired result.
\end{example}
\begin{example}
    Similarly, we can identify $K(\Z/2\Z,1)$ as $\RP^\infty$.
\end{example}

Since $\pi_1(K(A,n)) = 0$ for $n>1$, it follows that $K(A,n)$ is automatically
a simple space. This means that
$$[S^k,K(A,n)] = \pi_k(K(A,n)) = H^n(S^k,A).$$
In fact, a more general result is true:
\begin{theorem}[Brown representability]\label{brown-rep}
    If $X$ is a CW-complex, then $[X,K(A,n)] = H^n(X;A)$. 
\end{theorem}
We will not prove this here, but one can show this simply by showing that the
functor $[-,K(A,n)]$ satisfies the Eilenberg-Steenrod axioms. Somehow, these
Eilenberg-MacLane spaces $K(A,n)$ completely capture cohomology in dimension
$n$. 

If $X$ is a CW-complex, then we may assume that $X_{\leq n}$ is also a
CW-complex. (Otherwise, we can use cellular approximation and then kill
homotopy groups.) Let us assume that $X$ is path connected; then $X_{\leq 1} =
K(\pi_1(X),1)$. We may then form a (commuting) tower:
\begin{equation*}
    \xymatrix{
	& \vdots\ar[d] & \cdots\ar[l]\\
	& X_{\leq 3}\ar[d]& K(\pi_3(X),3)\ar[l]\\
	& X_{\leq 2}\ar[d]& K(\pi_2(X),2)\ar[l]\\
	X\ar[r]\ar[ur]\ar[uur]\ar[uuur]& X_{\leq 1}\ar@{=}[r] & K(\pi_1(X),1),
    }
\end{equation*}
since $K(\pi_n(X),n)\to X_{\leq n}\to X_{\leq n-1}$ is a fiber sequence.
This decomposition of $X$ is called the \emph{Postnikov tower} of $X$.

Denote by $X_{>n}$ the fiber of the map $X\to X_{\leq n}$ (for instance,
$X_{>1}$ is the universal cover of $X$); then, we have
\begin{equation*}
    \xymatrix{
	\cdots\ar[r]\ar[d] & \cdots\ar[r]\ar@{=}[d] & \vdots\ar[d] &
	\cdots\ar[l]\\
	X_{>3}\ar[r]\ar[d] & X\ar[r]\ar@{=}[d] & X_{\leq 3}\ar[d]&
	K(\pi_3(X),3)\ar[l]\\
	X_{>2}\ar[r]\ar[d] & X\ar[r]\ar@{=}[d] & X_{\leq 2}\ar[d]&
	K(\pi_2(X),2)\ar[l]\\
	X_{>1}\ar[r]\ar[d] & X\ar[r]\ar@{=}[d] & X_{\leq 1}\ar@{=}[r]\ar[d] &
	K(\pi_1(X),1)\\
	X\ar@{=}[r] & X\ar[r] & \ast
    }
\end{equation*}
The leftmost tower is called the \emph{Whitehead tower} of $X$, named after
George Whitehead.

I can take the fiber of $X_{>1}\to X$, and I get $K(\pi_1(X),0)$; more
generally, the fiber of $X_{>n} \to X_{>n-1}$ is $K(\pi_n(X),n-1)$. This yields
the following diagram:
\begin{equation*}
    \xymatrix{
	\cdots & \vdots\ar[d] & \vdots\ar@{=}[d] & \vdots\ar[d] & \cdots\\
	K(\pi_3(X),2)\ar[r] & X_{>3}\ar[r]\ar[d] & X\ar[r]\ar@{=}[d] & X_{\leq
	3}\ar[d]& K(\pi_3(X),3)\ar[l]\\
	K(\pi_2(X),1)\ar[r] & X_{>2}\ar[r]\ar[d] & X\ar[r]\ar@{=}[d] & X_{\leq
	2}\ar[d]& K(\pi_2(X),2)\ar[l]\\
	K(\pi_1(X),0)\ar[r] & X_{>1}\ar[r]\ar[d] & X\ar[r]\ar@{=}[d] & X_{\leq
	1}\ar@{=}[r]\ar[d] & K(\pi_1(X),1)\\
	& X\ar@{=}[r] & X\ar[r] & \ast
    }
\end{equation*}

We can construct Eilenberg-MacLane spaces as cellular complexes by attaching
cells to the sphere to kill its higher homotopy groups. The complexity of
homotopy groups, though, shows us that attaching cells to compute the
cohomology of Eilenberg-MacLane spaces is not feasible.
%These constructions go back to the 50's, and they had voluminous computations in low dimensions. One day in 1950, they got a postcard from Serre, who said, ``here's a computation you might be interested in: $H^{23}(K(\Z,14)) = ...$''. Of course, Serre and Cartan had a different approach, that was much more effective. They observed that the fact $\Omega K(\pi,n)\simeq K(\pi,n-1)$ wasn't perceived to be useful by Eilenberg and Maclane. They didn't think about fiber sequences. Serre and Cartan did this by means of a spectral sequence. We'll do that later in the course.
